String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-08-18 2023-1-19/sections/implementation.tex
setta2023

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH~\cite{Ostrich,atva2020}, resulting in a version of OSTRICH called $\ostrichrecl$.
Then we do experiments on two benchmark suites to evaluate the performance of $\ostrichrecl$. In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

@@

\vspace{-2mm}
\subsection{Benchmark suites}\label{sec:bench}

Our experiments utilize two benchmark suites, namely, RegCoL and AutomatArk. In total, there are 48,834 instances in the two benchmark suites. All the benchmark instances are in SMTLIB2 format.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite. These constraints are generated by extracting regular expressions with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. 
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraint $|x| > 10$ (where $x$ is a string variable). There are 8,215 instances in the AutomatArk suite here.
Note that the original AutomatArk benchmark suite in \cite{z3str3re} includes 19,979 instances in total, conjunctions of regular membership queries generated out of regular expressions from \cite{automatark}. 
%Only 8,751 ones that contain counting operators out of 19,979 instances.

Moreover, among the 48,834 RECL constraints in the RegCoL or AutomatArk suite, only 5\% of them include non-register-representable regexes.

%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
\hide{
We conducted a comparison on four sets of benchmarks based on regex with 
counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with the counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from the real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of the counting operator (the sum of upper bounds is greater than 50), and "small" contains remaining instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\vspace{-3mm}
\subsection{Experiment results}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5 
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3-Trau 
%(github commit 1628747) 
\cite{z3trau}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH\footnote{The name OSTRICH+ was used in~\cite{atva2020}, where nuXmv model checker was called to solve the nonemptiness of CEFA w.r.t. LIA. } 
%(github commit 8297d8d) 
\cite{atva2020}, on RegCoL and AutomatArk. All experiments are conducted on CentOS Stream release 8 with 12 Intel(R) Xeon(R) Platinum 8269CY 3.10GHz CPU cores and 190 GB memory. We use the Zaligvinder framework \cite{zaligvinder_2021} to execute the experiments, where the timeout period is set as 60 seconds per instance.

%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the size-reduction and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}

\begin{table}[ht]
\vspace{-3mm}
\begin{center}
  \import{tables}{table_regcol.tex}
\end{center}
  \caption{Experiment results on RegCol, with timeout = 60 seconds}
  \label{tab:results_regcol}
\vspace{-6mm}
\end{table}

The RegCol benchmark suite's experiment results are in Table~\ref{tab:results_regcol}. 
From Table~\ref{tab:results_regcol}, we can see that $\ostrichrecl$ solvers almost all 40,628 instances, except 106 of them, that is, it solves 40,522 instances correctly, which is %3,994/2,272/24,577/12,832/1,091/7,092 
3,893/2,081/20,483/10,200/1,082/5,505 more than the number of instances solved by CVC5/Z3seq/Z3-Trau/Z3str3/Z3str3RE/OSTRICH respectively. 
%    
Note that the solvers  Z3seq, Z3-Trau, Z3str3, and Z3str3RE have soundness errors, where the result of CVC5 is taken as the ground truth\footnote{Although it seems strange to use one solver as the ground truth, this is the practice for string constraint solving, since CVC5 is the most stable string solver at present.}. If CVC5 reports ``unknown'' or ``timeout'' in some instances, then no soundness errors will be reported in this instance. 
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver. More specifically, $\ostrichrecl$ spends 3.02 seconds per instance on average, while the fastest solver Z3str3RE spends 1.71 seconds per instance, and the other solvers spend at least 3.56 seconds per instance.  
Note that Z3-Trau reports ``unknown'' on 21,655 instances, which seems mainly attributed to the fact that it does not support \verb|re.diff|, the language difference operator of regular expressions.

\begin{table}[ht]
%\vspace{-1mm}
\begin{center}
  \import{tables}{table_automatark.tex}
\end{center}
  \caption{Experiment results on AutomatArk, with timeout = 60 seconds}
  \label{tab:results_automatark}
\vspace{-6mm}
\end{table}

The experiment results on the AutomatArk benchmark suite are put in Table~\ref{tab:results_automatark}. 
From Table~\ref{tab:results_automatark}, we can see that $\ostrichrecl$ solvers almost all 8,215 instances, except 39 of them, that is, it solves 8,176 instances correctly, which is %3,994/2,272/24,577/12,832/1,091/7,092 
101/191/4,094/2,632/9/1,587
more than the number of instances solved by CVC5/Z3seq/Z3-Trau/Z3str3/Z3str3RE/OSTRICH respectively. 
%    
Still, the soundness errors are reported by taking the results of CVC5 as the ground truth.
%Note that Z3-Trau has soundness errors, where the result of CVC5 is taken as the ground truth. If CVC5 reports ``unknown'' or ``timeout'' in some instances, then no soundness errors will be reported in this instance. 
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ spends 2.49 seconds per instance on average. 
%while the fastest solver Z3str3RE spends 1.71 seconds per instance, and the other solvers spend at least 3.49 seconds per instance.  
%Note that Z3-Trau reports ``unknown'' on 2,552 instances, which seems to be mainly attributed to the fact that it does not support \verb|re.diff|, the language difference operator of regular expressions.

In summary, $\ostrichrecl$ solves 48,698 out of 48,834 instances, 1,091 more than Z3str3RE, the state-of-the-art best solver on RECL constraints. Moreover, the speed of $\ostrichrecl$ is the second fastest and comparable to that of Z3str3RE. (See Table~\ref{tab:results_summary}.)

\begin{table}[ht]
\vspace{-3mm}
\begin{center}
  \import{tables}{table_summary.tex}
\end{center}
  \caption{Experiment results: A Summary}
  \label{tab:results_summary}
\vspace{-6mm}
\end{table}

>>>>

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH~\cite{Ostrich,atva2020}, resulting in a version of OSTRICH called $\ostrichrecl$.
Then we do experiments on two benchmark suites to evaluate the performance of $\ostrichrecl$, and carry out detailed experiments including evaluation of size-reduction techniques and the experiment of large-counting benchmark. In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

@@

\vspace{-3mm}
\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}
\vspace{-2mm}

Our experiments utilize two benchmark suites, namely, RegCoL and AutomatArk. In total, there are 48,843 instances in the two benchmark suites. All the benchmark instances are in SMTLIB2 format.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite. These constraints are generated by extracting regular expressions with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form
$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$
where $\overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. The artificial construction is meaningful because the special characters are offensive. For instance, they are widely used in client-side attacks like XSS\cite{malware_detection_3_kudzu}\cite{CCH+18}.
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraint $|x| > 10$ (where $x$ is a string variable). There are 8,215 instances in the AutomatArk suite here.
Note that the original AutomatArk benchmark suite in \cite{z3str3re} includes 19,979 instances in total, conjunctions of regular membership queries generated out of regular expressions from \cite{automatark}.
%Only 8,751 ones that contain counting operators out of 19,979 instances. 
Moreover, only 5\% of them include non-register-representable regexes among the 48,843 instances.
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

\medskip
\noindent
\emph{Experiment setup.}
All experiments are conducted on CentOS Stream release 8 with 4 Intel(R) Xeon(R) Platinum 8269CY 3.10GHz CPU cores and 190 GB memory. We use the \textsc{zaligvinder} framework \cite{zaligvinder_2021} to execute the experiments, with a timeout of 60s for each instance.
\hide{
  We conducted a comparison on four sets of benchmarks based on regex with
  counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with the counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from the real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of the counting operator (the sum of upper bounds is greater than 50), and "small" contains remaining instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\vspace{-3mm}
\subsection{Overall Experiment Results}
\vspace{-2mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
%(github commit 8297d8d) 
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
% Z3-trau \cite{z3trau} is not evaluated because it does not support \verb|re.diff|, the language difference operator of regular expressions.  
%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the size-reduction and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}
\vspace{-5mm}
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall Experiment results,\\ with timeout = 60 seconds}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Simplification technique evaluation, with timeout = 60 seconds}
      \label{tab:results_simp}
  \end{subtable}
  \caption{Experiments on RegCoL and AutomatArk benchmark suites}
\end{table}
\vspace{-9mm}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

The overall experiment results are in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{At first we used the majority vote between all results of the solvers as the ground truth, but solvers of z3 family outputed some wrong results(verified by us manually), causing the ground truth to be incorrect.} and the soundness error also be counted as unknown. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, our solver $\ostrichrecl$ is the second fastest solver and its average time on each instance is closed to the fastest solver Z3str3RE (\textbf{1.93}/1.62 s).
\vspace{-3mm}
\subsection{Evaluation of Simplification Techniques}
\vspace{-2mm}

We evaluate the simplification techniques by comparing the performance of $\ostrichrecl$ with and without the simplification techniques. Moreover, we also encode the $\cefadec$ problem to infinite state transition system and solve it using \textsc{nuXmv}\cite{nuxmv}. 
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_simp.tex}
%   \end{center}
%   \caption{Simplification technique evaluation, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}

The evaluation results of simplification techniques are in Table~\ref*{tab:results_simp}. We can see that compared to $\ostrichrecl$ without simplification techniques, $\ostrichrecl$ with simplification techniques solves \textbf{1,503 more} instances and is \textbf{2.21$\times$} faster. Moreover, compared with the techniques using \textsc{nuxmv}, $\ostrichrecl$ with simplification techniques solves \textbf{1,798 more} instances and is \textbf{3.13$\times$} faster.

\vspace{-3mm}
\subsection{Large-Counting Experiment Results}
The distribution of upper bound in RegCoL and AutomatArk suites is showed in Fig~\ref{fig:count_distri}, whose horizontal axis is the value of upper bound and vertical axis is the number of occurrences of it. The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). Although the upper bounds of most counting operators are less than 50, there are still about two thousands of them are greater than 50. We extract these instances with counting upper bound greater than 50 (1,969 instances in total). Moreover, We change the length constraint to $|x| > 200$ (which is $|x| > 10$ in overall experiment), to make the number of $sat$ result more closed to that of $unsat$.
\vspace{-5mm}
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of counting operators in two benchmark suites.}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Large counting experiment, \\with timeout = 60 seconds}
    \label{fig:table_large_count}
  \end{subfigure}
  \caption{Analysis of dataset and large counting experiment}
\end{figure}
\vspace{-7mm}

% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_large_count.tex}
%   \end{center}
%   \caption{Large counting experiment, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}

We evaluate the performance of $\ostrichrecl$ on these instances. The results are in Fig~\ref{fig:table_large_count}. We can see that $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, \newline
$\ostrichrecl$ is the fastest solver, who is \textbf{ 6.79/2.88/2.61/5.27/3.95$\times$ } faster than CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
\vspace{-2mm}

% The results are in Table~\ref{tab:results_simplification}. We can see that the simplification technique improves the performance of $\ostrichrecl$ significantly. It solves 1,111 more instances and is 1.62 times faster than $\ostrichrecl$ without the simplification technique.

% \begin{table}[ht]
% %\vspace{-1mm}
% \begin{center}
%   \import{tables}{table_automatark.tex}
% \end{center}
%   \caption{Experiment results on AutomatArk, with timeout = 60 seconds}
%   \label{tab:results_automatark}
% \vspace{-6mm}
% \end{table}

% The experiment results on the AutomatArk benchmark suite are put in Table~\ref{tab:results_automatark}. 
% From Table~\ref{tab:results_automatark}, we can see that $\ostrichrecl$ solvers almost all 8,215 instances, except 39 of them, that is, it solves 8,176 instances correctly, which is %3,994/2,272/24,577/12,832/1,091/7,092 
% 101/191/4,094/2,632/9/1,587
% more than the number of instances solved by CVC5/Z3seq/Z3-Trau/Z3str3/Z3str3RE/OSTRICH respectively. 
% %    
% Still, the soundness errors are reported by taking the results of CVC5 as the ground truth.
%Note that Z3-Trau has soundness errors, where the result of CVC5 is taken as the ground truth. If CVC5 reports ``unknown'' or ``timeout'' in some instances, then no soundness errors will be reported in this instance. 
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
% Moreover, $\ostrichrecl$ spends 2.49 seconds per instance on average. 
%while the fastest solver Z3str3RE spends 1.71 seconds per instance, and the other solvers spend at least 3.49 seconds per instance.  
%Note that Z3-Trau reports ``unknown'' on 2,552 instances, which seems to be mainly attributed to the fact that it does not support \verb|re.diff|, the language difference operator of regular expressions.

% In summary, $\ostrichrecl$ solves 48,698 out of 48,843 instances, 1,091 more than Z3str3RE, the state-of-the-art best solver on RECL constraints. Moreover, the speed of $\ostrichrecl$ is the second fastest and comparable to that of Z3str3RE. (See Table~\ref{tab:results_summary}.)

% \begin{table}[ht]
% \vspace{-3mm}
% \begin{center}
%   \import{tables}{table_summary.tex}
% \end{center}
%   \caption{Experiment results: A Summary}
%   \label{tab:results_summary}
% \vspace{-6mm}
% \end{table}
